### 2015-05-10 Report/article_2.tex
Added Cleaned Programs & Report

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Journal Article
% LaTeX Template
% Version 1.3 (9/9/13)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Frits Wenneker (http://www.howtotex.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[twoside]{article}

\usepackage{lipsum} % Package to generate dummy text throughout this template

\usepackage[sc]{mathpazo} % Use the Palatino font
\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
\linespread{1.05} % Line spacing - Palatino needs more space between lines
\usepackage{microtype} % Slightly tweak font spacing for aesthetics
\usepackage{amsmath}

\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry} % Document margins
\usepackage{multicol} % Used for the two-column layout of the document
\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
\usepackage{booktabs} % Horizontal rules in tables
\usepackage{float} % Required for tables and figures in the multi-column environment - they need to be placed in specific locations with the [H] (e.g. \begin{table}[H])
\usepackage{hyperref} % For hyperlinks in the PDF

\usepackage{lettrine} % The lettrine is the first enlarged letter at the beginning of the text
\usepackage{paralist} % Used for the compactitem environment which makes bullet points with less space between them
\usepackage{graphicx}
\usepackage{abstract} % Allows abstract customization
\renewcommand{\abstractnamefont}{\normalfont\bfseries} % Set the "Abstract" text to bold
\renewcommand{\abstracttextfont}{\normalfont\small\itshape} % Set the abstract itself to small italic text

\usepackage{titlesec} % Allows customization of titles
\renewcommand\thesection{\Roman{section}} % Roman numerals for the sections
\renewcommand\thesubsection{\Roman{subsection}} % Roman numerals for subsections
\titleformat{\section}[block]{\large\scshape\centering}{\thesection.}{1em}{} % Change the look of the section titles
\titleformat{\subsection}[block]{}{\thesubsection.}{1em}{} % Change the look of the section titles

\usepackage{fancyhdr} % Headers and footers
\pagestyle{fancy} % All pages have headers and footers
\fancyhead{} % Blank out the default header
\fancyfoot{} % Blank out the default footer
%\fancyhead[C]{Running title $\bullet$ November 2012 $\bullet$ Vol. XXI, No. 1} % Custom header text
\fancyfoot[RO,LE]{\thepage} % Custom footer text

%----------------------------------------------------------------------------------------
%	TITLE SECTION
%----------------------------------------------------------------------------------------

\title{\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{Understanding Driver Behavior: Studies on NYC Taxi Data}} % Article title

\author{
\large
\textsc{John Smith}\thanks{A thank you or further information}\\[2mm] % Your name
\normalsize University of California \\ % Your institution
\normalsize \href{mailto:john@smith.com}{john@smith.com} % Your email address
\vspace{-5mm}
}

\author{
\large
\textsc{Daniel Abramson}\\[2mm] % Your name
\normalsize New York University \\ % Your institution
\normalsize \href{mailto:dna237@nyu.edu}{dna237@nyu.edu} % Your email address
\vspace{-5mm}
\and
\textsc{Yoon Kim}\\[2mm] % Your name
\normalsize New York University \\ % Your institution
\normalsize \href{mailto:yhk255@nyu.edu}{yhk255@nyu.edu} % Your email address
\vspace{-5mm}
\and
\textsc{Sylvia Zhao}\\[2mm] % Your name
\normalsize New York University \\ % Your institution
\normalsize \href{mailto:ssz8@cornell.edu}{ssz8@cornell.edu} % Your email address
\vspace{-5mm}
}
\date{}

%----------------------------------------------------------------------------------------

\begin{document}

\maketitle % Insert title

\thispagestyle{fancy} % All pages have headers and footers

%----------------------------------------------------------------------------------------
%	ABSTRACT
%----------------------------------------------------------------------------------------

\begin{abstract}

\noindent Using all data on NYC taxi trips and fares from 2010 and 2013 we perform a series of studies to understand motivating factors behind driver behavior as well as outcomes from those incentives. In order to analyze this large dataset ($\approx$14GB), we employ a number of big data processing technologies that work on the Hadoop framework: Hive, Spark, and MapReduce programs.  We present our findings, issues encountered, and experimental setup. Of these findings, the most notable include: there are outlier drivers who are on average more productive per day, underperforming drivers generally do not improve (in relation to their peers), the sensitivity of fares to environment variables are seemingly invariant, drivers can expect better tips from cars with certain numbers of passengers, and previous tips affect driver behavior for the next immediate trip.

\end{abstract}

%----------------------------------------------------------------------------------------
%	ARTICLE CONTENTS
%----------------------------------------------------------------------------------------

\begin{multicols}{2} % Two-column layout throughout the main article text

\section{Introduction}

\lettrine[nindent=0em,lines=3]{D} river behavior presents itself as an interesting facet of the NYC taxi data to study through a number of different lenses. As a passenger, driver behavior affects the quality of the trip, cost of the trip, as well as availability of the taxi (in order to even take the trip in the first place). From the perspective from the taxi company, driver behavior affects revenues and profitability. As such, the taxi company may, through a better understanding of such behavior, change its incentive structure to encourage certain behaviors over others (now that it knows more about the current environment).

We ask a series of questions that are pertinent to both lenses. 
\begin{enumerate}
\item Are there productivity differences among drivers? Do some drivers consistently make more revenue per unit time worked than others? 
\item Can we predict the sensitivities of taxi fares to a number of different environmental variables that would affect driver incentives to work? These include hours of the day, days of the week, and seasonal factors.
\item Using this model, can we identify drivers performance differences? How do these relate to characteristic variables of the trip? Are tips sensitive to these outliers in performance/are passengers aware of these performance differences?
\item Do these sensitivities of taxi fares to environmental variables change over time?
\item Do previous earnings affect driver behavior? Is there some sort of 'hot-hand' with respect to tips? If a driver receives an above-median tip does this affect his behavior on the next ride? Is he motivated to continue good performance or does he care less? Is there no change in behavior?

\end{enumerate}

%------------------------------------------------

\section{Preprocessing Taxi Data Inputs}

All of the analyses above used the trips and fares data for NYC in 2010. Comparisons over time used the data for 2013 as well. This data is all publicly available. \cite{NYC_DATA}

As a starting point for our analyses we joined these two datasets. In order to perform this inner join, we used MrJob, a python tool for using Hadoop-Streaming with a programming model that more closely follows Java.\cite{MRJOB}  For instance, instead of writing a reducer that iterates through each line to collect values for a given key, MrJob provides the reducer with a key and list of values.

We used AWS to perform this join, first transferring all input files to S3. We then used 4 m3.xlarge machines to perform the join. This took about 2 hours for each. There were 7 reducers. Mapper numbers were set to defaults.

%------------------------------------------------

\section{Investigating Question 1: Driver Productivity Differences}

\subsection{Experimental Setup}
In answering this question, we performed a simple MapReduce analysis via MrJob. For each trip point, we had the mapper output the driver, date, revenue, and minutes worked. We then used a combiner to sum the last 2 quantities for a given driver-date pair.  Lastly, the producer output the daily average of the last 2 quantities for a given driver. Nonsense points where either fares were negative or there were more minutes than in a workday were removed.

The experimental setup of the hadoop environment is the same as that described in Section II. The runtime is similarly 2 hours. After running this job, aggregation of the results was performed locally via a single python script.

\subsection{Results}
\includegraphics[scale=.35]{Minutes_v_Rev.png}

Though there is nothing surprising about a positive linear relationship between Daily Revenue and Minutes Worked per day, we find the outliers to be interesting. There seems to be a larger spread of outliers below the line than above. However, analyzing the outliers above the line, some seem to work very few minutes an contribute a lot to revenue. Perhaps these drivers are not properly incentivized. It would appear they have much downtime during the day.

\section{Investigating Question 2: Fare Sensitivities}

\subsection{Experimental Setup}

We leveraged Spark to perform this analysis. Specifically, we performed a number of regressions on the taxi data from 2010 and 2013 (as well as the combined dataset). We set up a regression model of the form:

\begin{multline*}
\text{Fare}[i] = \alpha + \\
	\beta \cdot [\text{distance}[i], \{H[i]\}, \{D[i]\}, \{M[i]\},\{HD[i]\}] +\\
	 \epsilon[i]
\end{multline*}

\noindent with an appropriate number of dummies for the hour of the day, day of the week, month, and hour-day interactions. We dropped nonsense points and outliers: fares less than 0, distances less than 0 or bigger than 100 miles.

We first made the non binary features standard normal. We then used Spark to fit this linear model. Spark forms the Mean-Squared-Error loss function and performs stochastic gradient descent updates over the inputs to adjust the weights to minimize the function. We utilized 4 m3.2xlarge machines to perform the analysis on AWS. This took about 40 hours (using 100 iterations for each fitting of the model). We then saved the predictions, actual values, and trip datapoint characteristics as a new dataset.

Additionally, we approximated the variance of $\hat{\beta}$. We found this to be an interesting problem as an exact result would require quadratic time, while an approximate one, like jackknifing would also be too computationally intensive (and of the same complexity though not assuming a functional form like homoskedasticity).  We devised an approximation by randomly sampling 300,000 points (call this $\tilde{X}$) without replacement from our dataset. We then approximated $Var(\hat{\beta})$ by $\text{MSE}*(\frac{N}{300000}\tilde{X}^T\tilde{X})^{-1}$. We used Spark for this sampling procedure to produce our homeskedastic standard errors. 
\subsection{Results}

The beta values are presented below in order of largest value to smallest value. A flag is included for significance.

Below, we present them in order of most to least significant (in terms of the t-statistic). We include a flag for significance.

Please see appendix for exact betas and standard errors.

We see these as an interesting factor in driver behavior because...

\section{Investigating Question 3: Driver Performance v Fare Sensitivities}
YOOOOOON
\subsection{Experimental Setup}
\subsection{Results}

\section{Investigating Question 4: Time v Fare Sensitivities}

\subsection{Experimental Setup}
In order to test time invariance we fit regressions on both the 2010 and 2013 joined taxi data (separately and together). Using Spark, we jointly test whether the regression coefficients are the same as well as perform a test per coefficient for equality.

The first analysis is conducted via a Chow Test.  The null hypothesis of the Chow Test is that $\beta_{2010} = \beta_{2013}$.  The Chow Statistic is formed by

$$\frac{(SSR_{comb} - (SSR_{2010} + SSR_{2013}))/k}{(SSR_{2010} + SSR_{2013})/(N_{2010} + N_{2013}-2k)}$$

\noindent which is distributed $F(k, N_{2010} + N_{2013}-2k)$. \cite{CHOW}

We used Spark to calculate the Sum-of-Squared-Residual terms from the fitted models as well as the counts. This was included in the original regression code (so as to avoid loading the data twice).

Secondly, we performed a Z-Test on each individual coefficient. Here

$$ Z_i = \frac{\hat{\beta}_{2010}[i] - \hat{\beta}_{2013}[i]}{\sqrt(SE(\hat{\beta}_{2010}[i]) + SE(\hat{\beta}_{2013}[i]))}.$$

\subsection{Results}

For the Chow statistic we find

For the individual Z-tests we present the figure below in order of absolute value $Z_i$.

Discussion

\section{Investigating Question 5: Previous Earnings v Behavior}
\subsection{Experimental Setup}
For this analysis, we return to MrJob. Here our map task outputs $\text{[driver, (pickup day, tip percentage)]}$
for every entry in our joined taxi data. We then implement a rather complicated reducer.  For every driver, the reducer computes the tip percentage median per day. This is used to control for possible trends in the data (we assume a day is small enough to not contain an economic trend). We then iterate through all of the tip percentages each day and count the number of times a tip above the daily median is followed by another vs one below the daily median. These are `above-above' vs `above-below'. By definition, half the tips must be above the median and half below. Hence, we find the frequency of `above-above' vs `above-below' to be a good test for a behavioral factor. For each driver we output the tip statistic,

$$\frac{Count(`above-above')}{Count(`above-above') + Count(`above-below')}.$$

We then aggregate these results and look at the distribution over drivers. The running time and experimental setup is nearly identical to our the MR jobs. 
\subsection{Results}
The resulting distribution is depicted below. 
\includegraphics[scale=.35]{hothand.png}

For various bin sizes, we find a similarly bimodal result. The red line represents the resulting normal distribution given our input data. Here, we see that the distribution is definitely not normal. However, behavior is on both ends of the spectrum. Moreover, we are unsure if it is exactly the behavior we thought we were measuring in the first place. We assumed that a good tip would improve driver moral and lead him to try harder on the next trip.  Then he would be more likely to get a larger tip.  Here we also see that for many drivers a good tip is more likely followed by a bad tip.  Either a good tip leads to laziness or we need more controls. Perhaps a time-of-day control could improve the experiment assuming that tipping trends change in morning vs afternoon vs night. This may be a reasonable assumption. We did not have sufficient funds to continue this analysis on AWS.

%------------------------------------------------

\section{Concluding Remarks}

%----------------------------------------------------------------------------------------
%	REFERENCE LIST
%----------------------------------------------------------------------------------------

\begin{thebibliography}{99} % Bibliography - this is intentionally simple in this template

\bibitem[Figueredo and Wolf, 2009]{Figueredo:2009dg}
Figueredo, A.~J. and Wolf, P. S.~A. (2009).
\newblock Assortative pairing and life history strategy - a cross-cultural
  study.
\newblock {\em Human Nature}, 20:317--330.

\end{thebibliography}

%----------------------------------------------------------------------------------------

\section*{Appendices}

\subsection{Running the Code}

Our code is available at \\

\noindent https://github.com/ssz225/bigdata\_final.\\

Each component has a corresponding question folder with instructions as to how to run the code. We have excluded all of our AWS access information from the code and put in dummy values.  The joined input data is available at \\

\noindent https://s3.amazonaws.com/trip-fare-join/2010/part-0000X
\\
\noindent https://s3.amazonaws.com/trip-fare-join/2013/part-0000X
\\

\noindent where X ranges from 0 to 6.

\subsection{Individual Contributions}

\end{multicols}

\end{document}

### 2015-05-10 Q1 - Driver Productivity/histo.py

@@ -2,6 +2,11 @@ import glob
 import numpy as np
 import matplotlib.pyplot as plt
 import matplotlib.mlab as mlab
+from scipy.stats import multivariate_normal
+from mpl_toolkits.mplot3d import Axes3D
+from matplotlib import cm
+from matplotlib.ticker import LinearLocator, FormatStrFormatter
+
 
 in_files = glob.glob('output-2010/p*')
 
@@ -29,4 +34,5 @@ plt.grid(True)
 plt.xlabel('Minutes Worked/Day')
 plt.ylabel('Daily Revenue')
 plt.savefig('Minutes_v_Rev.png', dpi=400)
+plt.close()
 

### 2015-05-10 Q2&4 - Fare Sensitivities/spark_reg_local.py

@@ -1,0 +1,163 @@
+from pyspark import SparkContext
+from pyspark.mllib.regression import LabeledPoint, LinearRegressionWithSGD
+from numpy import array
+from datetime import datetime
+import math
+from pyspark.mllib.util import MLUtils
+from pyspark.mllib.linalg import Vectors
+import numpy as np
+from numpy.linalg import inv
+from pyspark.mllib.stat import Statistics
+
+
+n_scaled = 1
+avg_dist_fill = 2.6
+n_samps = 300000
+
+in_loc = "FILL IN"
+out_loc = "FILL IN"
+
+
+def get_total_seconds(td): 
+	return (td.microseconds + (td.seconds + td.days * 24 * 3600) * 1e6) / 1e6
+
+fmt = '%Y-%m-%d %H:%M:%S'
+# Load and parse the data
+def parsePoint(line):
+	items = line.strip('"').split('"\t"')
+	items = items[0].split(',') + items[1].split(',')
+	p_up = datetime.strptime(items[3], fmt)
+	d_time = datetime.strptime(items[6], fmt)
+	#minutes_worked = get_total_seconds(d_time - p_up)/60.0
+	fare_amount = float(items[-6])
+	try:
+		distance = float(items[9])
+	except:
+		distance = avg_dist_fill
+	hour_of_day = p_up.hour
+	day_of_week = p_up.weekday()
+	hour_dummies = [0]*23
+	try:
+		hour_dummies[hour_of_day] = 1
+	except:
+		pass
+	day_dummies = [0]*6
+	try:
+		day_dummies[day_of_week] = 1
+	except:
+		pass
+	month = p_up.month-1
+	month_dummies = [0]*11
+	try:
+		month_dummies[month] = 1
+	except:
+		pass
+	interactions = np.zeros((7,24))
+	interactions[day_of_week,hour_of_day] = 1
+	interactions = interactions.flatten().tolist()[:-1] #correct perfect collinearity
+	X = [distance]
+	X = X +  hour_dummies+  day_dummies + month_dummies + interactions
+	#filter junk
+	if fare_amount>0 and distance<100 and distance>0:
+		return LabeledPoint(fare_amount, X)
+
+
+def parseKeepD(line):
+	items = line.strip('"').split('"\t"')
[…]

### 2015-05-10 Q4-regression/reg_control.py

@@ -1,31 +1,0 @@
-import sys
-import string
-import os
-import numpy as np
-import subprocess
-import re
-
-weights = [10,10]
-
-
-def make_run(weights):
-	part_1 = 'python regression.py -r local --cmdenv weights='
-	weight_str = ','.join([str(w) for w in weights])
-	part_2 = ' test_inputs/'
-	return part_1 + weight_str + part_2
-
-diff = 100
-tolerance = 10**-5
-
-while diff > tolerance:
-	run_str = make_run(weights)
-	last_weights = np.array(weights)
-	p = subprocess.Popen(run_str, shell=True, stdout=subprocess.PIPE, stderr=subprocess.STDOUT)
-	x= p.stdout.readlines()
-	for i in range(0,len(x)):
-		if re.match('Streaming final output from', x[i]):
-			weights = [float(y) for y in x[i+1].strip('1\t\'[\]\n').split(',')]
-	#l2 
-	print weights
-	diff = np.linalg.norm(np.array(weights) - last_weights, 2)
-

### 2015-05-10 Q4-regression/regression.py

@@ -1,62 +1,0 @@
-import sys
-import string
-import os
-import numpy as np
-from datetime import datetime
-from sklearn import linear_model
-
-fmt = '%Y-%m-%d %H:%M:%S'
-from mrjob.job import MRJob
-
-
-
-class Regressor(MRJob):
-	
-	def mapper_init(self):
-		self.fitter = linear_model.SGDRegressor(loss='squared_loss', penalty='none', fit_intercept=True, random_state=None)
-		weights = [float(x) for x in os.environ['weights'].split(',')]
-		if weights is not None:
-			self.fitter.coef_ = np.array(weights[1:])
-			self.fitter.intercept_ = np.array([weights[0]])
-
-
-	def mapper(self, _, line):
-		#this should create a single input
-		items = line.strip('"').split('"\t"')
-		items = items[0].split(',') + items[1].split(',')
-
-		p_up = datetime.strptime(items[3], fmt)
-		d_time = datetime.strptime(items[6], fmt)
-
-
-		minutes_worked = (d_time - p_up).total_seconds()/60.0
-
-		#our y var
-		fare_amount = float(items[-6])
-		if fare_amount>0:
-			#some controls
-			day_of_week_flag = p_up.weekday()
-			hour = p_up.hour
-			#some time buckets (for hours in day)
-
-			#fit and update the fitter
-			X = np.array([[minutes_worked]])
-
-			self.fitter = self.fitter.partial_fit(X, [fare_amount])
-
-
-	def mapper_final(self):
-		full_weights = list(self.fitter.intercept_) + list(self.fitter.coef_ )
-		yield (1, full_weights)
-
-
-	def reducer(self, key, values):
-		v= [x for x in  values]
-		weight_arr = list(np.mean(np.array(v), axis=0))
-		yield(1, weight_arr)
-
-
-
-
-if __name__ == '__main__':
-    Regressor.run()

### 2015-05-10 Q4-regression/spark_reg.py

@@ -1,31 +1,0 @@
-from pyspark.mllib.regression import LabeledPoint, LinearRegressionWithSGD
-from numpy import array
-from datetime import datetime
-
-def get_total_seconds(td): 
-	return (td.microseconds + (td.seconds + td.days * 24 * 3600) * 1e6) / 1e6
-
-fmt = '%Y-%m-%d %H:%M:%S'
-# Load and parse the data
-def parsePoint(line):
-	items = line.strip('"').split('"\t"')
-	items = items[0].split(',') + items[1].split(',')
-	p_up = datetime.strptime(items[3], fmt)
-	d_time = datetime.strptime(items[6], fmt)
-	minutes_worked = get_total_seconds(d_time - p_up)/60.0
-	fare_amount = float(items[-6])
-	X = [minutes_worked]
-	return LabeledPoint(fare_amount, X)
-
-data = sc.textFile("s3n://AWS_ACCESS_KEY_ID:AWS_SECRET_KEY@trip-fare-join/2010/")
-parsedData = data.map(parsePoint)
-
-# Build the model
-model = LinearRegressionWithSGD.train(parsedData)
-
-# Evaluate the model on training data
-valuesAndPreds = parsedData.map(lambda p: (p.label, model.predict(p.features)))
-MSE = valuesAndPreds.map(lambda (v, p): (v - p)**2).reduce(lambda x, y: x + y) / valuesAndPreds.count()
-print("Mean Squared Error = " + str(MSE))
-
-

### 2015-05-10 Q5 - Hot Hand/histo.py

@@ -24,7 +24,7 @@ l = plt.plot(bins, y, 'r--', linewidth=1)
 
 plt.xlabel('P(tip(t+1)>median) | tip(t)>median)')
 plt.ylabel('Scaled Frequency')
-plt.title(r'Are Good tips Followed by Good Tips?')
+plt.title(r'Are Good Tips Followed by Good Tips?')
 plt.axis([.1, .9,0,10])
 plt.grid(True)
 

### 2015-05-10 QX - Reg Errors/histos.py

@@ -1,0 +1,50 @@
+import glob
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.mlab as mlab
+import json
+
+f_2010 = glob.glob('outputs/2010/p*')
+
+f_2013 = glob.glob('outputs/2013/p*')
+
+def file_loop(f):
+    lines = [line.strip().split('\t') for line in open(f)]
+    vals = [(l[0],json.loads(l[1])[0],json.loads(l[1])[1] )for l in lines]
+    return vals
+
+v_2010 = []
+for f in f_2010:
+    v_2010.extend(file_loop(f))
+
+
+v_2013 = []
+for f in f_2013:
+    v_2013.extend(file_loop(f))
+
+def make_graph(values, label):
+	n, bins, patches = plt.hist(values, 35, normed=1, stacked=1,facecolor='green', alpha=0.75)
+	# add a 'best fit' line (assuming normality)
+	y = mlab.normpdf( bins, np.mean(values), np.std(values))
+	l = plt.plot(bins, y, 'r--', linewidth=1)
+
+	plt.xlabel(label)
+	plt.ylabel('Scaled Frequency')
+	plt.title(r'Distribition of ?' + label)
+	#plt.axis([.1, .9,0,10])
+	plt.grid(True)
+
+	plt.savefig(label + '.png', dpi=400)
+	plt.close()
+
+#strip teh very strange outliers
+avg_err_10 = [v[1] for v in v_2010 if (x<1 and x>-1)]
+abs_err_10 = [v[2] for v in v_2010 if (x<50 and x>50)]
+
+avg_err_13 = [v[1] for v in v_2013 if (x<1 and x>-1)]
+abs_err_13 = [v[2] for v in v_2010 if (x<50 and x>50)]
+
+make_graph(avg_err_10, 'Average Errors 2010')
+make_graph(abs_err_10, 'Absolute Errors 2010')
+make_graph(avg_err_13, 'Average Errors 2013')
+make_graph(avg_err_13, 'Absolute Errors 2013')

### 2015-05-10 QX - Reg Errors/reg_errors.py

@@ -1,0 +1,36 @@
+import sys
+import string
+import os
+
+
+from mrjob.job import MRJob
+
+
+class RegErrors(MRJob):
+
+	def mapper(self, _, line):
+		x = line.strip('"(').split(",")[:4]
+		actual,pred = [float(a) for a in x[:2]]
+		m1 = x[2].strip('"[').strip('" [').strip('\'')
+		m2 = x[3].strip('"[').strip('" [').strip('\'')
+		driver = m1 + "_" + m2
+		yield (driver, [actual,pred])
+
+
+	def reducer(self, key, values):
+		vals = [v for v in values]
+		percs = []
+		absos = []
+		for v in vals:
+			try:
+				percs.append(float(v[0] - v[1]) / v[1])
+			except:
+				pass
+			try:
+				absos.append(float(v[0] - v[1]))
+			except:
+				pass
+		yield (key, [float(sum(percs))/len(percs), float(sum(absos))/len(absos)])
+
+if __name__ == '__main__':
+    RegErrors.run()

### 2015-05-10 geocode.py

@@ -1,15 +1,0 @@
-import cPickle as pickle
-import numpy as np
-m = pickle.load(open("zip_grid.p","rb"))
-
-def return_zip(m, lat, lon, nw_lat=40.917, nw_lon=-74.260, steps=0.002):
-    """
-    Returns zip code based on a grid search
-    """
-    lat, lon = float(lat), float(lon)     
-    i, j = int((nw_lat - lat)/steps), int((lon - nw_lon)/steps)
-    try:
-        return m[i][j]
-    except:
-        return -1
-        

### 2015-05-11 Report/article_2.tex
Updated Report with Yoon's parts

@@
 These include hours of the day, days of the week, and seasonal factors.
-\item Using this model, can we identify drivers performance differences?
-How do these relate to characteristic variables of the trip?
-Are tips sensitive to these outliers in performance/are passengers aware of these performance differences?
-\item Do these sensitivities of taxi fares to environmental variables change over time?
+\item Using this model, can we identify ``outlier" drivers whose actual fare exceeds the predicted fare (i.e. those that are overcharging)? How is tip behavior affected by such things?
 \item Do previous earnings affect driver behavior?
@@
 Is there no change in behavior?
+\item How is tip affected by the income of the passengers, as measured by the median household income of the zip code in which the passenger is dropped off?
 
@@
 \noindent with an appropriate number of dummies for the hour of the day, day of the week, month, and hour-day interactions.
-We dropped nonsense points and outliers: fares less than 0, distances less than 0 or bigger than 100 miles.
 
@@
 
-Additionally, we approximated the variance of $\hat{\beta}$.
+Additionally, we approximated the standard error of $\beta$.
 We found this to be an interesting problem as an exact result would require quadratic time, while an approximate one, like jackknifing would also be too computationally intensive (and of the same complexity though not assuming a functional form like homoskedasticity).
 We devised an approximation by randomly sampling 300,000 points (call this $\tilde{X}$) without replacement from our dataset.
-We then approximated $Var(\hat{\beta})$ by $\text{MSE}*(\frac{N}{300000}\tilde{X}^T\tilde{X})^{-1}$.
+We then approximated $Var(\beta)$ by $\text{MSE}*(\frac{N}{300000}\tilde{X}^T\tilde{X})^{-1}$.
 We used Spark for this sampling procedure to produce our homeskedastic standard errors.
@@
 \section{Investigating Question 3: Driver Performance v Fare Sensitivities}
-YOOOOOON
-\subsection{Experimental Setup}
-\subsection{Results}
+Based on the regression output from the previous section, we plot the residual (i.e. actual fare minus predicted fare) for each trip and bin them into histograms.  
+\includegraphics[scale=.35]{figure_1.png}
+While much the error may be coming from the fact that our regression model is very simple (for example, it doesn't take into account the fact that a distance of 1 mile means very different things in midtown versus in suburbs of Queens), there does seem to be a cohort of trips in which the customer is overcharged.
+\\
+Next we investigate whether customers who are overcharged (i.e.
+have high residuals) give less tip.
+Below is a scatter plot of residuals (x-axis) versus tip percentage (y-axis).
+While the relationship is far from clear, there does seem to be a slight negative relationship between tip percentage and residual.
+\includegraphics[scale=.35]{figure_2.png}
 
-\section{Investigating Question 4: Time v Fare Sensitivities}
-
-\subsection{Experimental Setup}
-In order to test time invariance we fit regressions on both the 2010 and 2013 joined taxi data (separately and together).
-Using Spark, we jointly test whether the regression coefficients are the same as well as perform a test per coefficient for equality.
-
-The first analysis is conducted via a Chow Test.  The null hypothesis of the Chow Test is that $\beta_{2010} = \beta_{2013}$.  The Chow Statistic is formed by
-
-$$\frac{(SSR_{comb} - (SSR_{2010} + SSR_{2013}))/k}{(SSR_{2010} + SSR_{2013})/(N_{2010} + N_{2013}-2k)}$$
-
-\noindent which is distributed $F(k, N_{2010} + N_{2013}-2k)$. \cite{CHOW}
-
-We used Spark to calculate the Sum-of-Squared-Residual terms from the fitted models as well as the counts. This was included in the original regression code (so as to avoid loading the data twice).
-
-Secondly, we performed a Z-Test on each individual coefficient. Here
-
-$$ Z_i = \frac{\hat{\beta}_{2010}[i] - \hat{\beta}_{2013}[i]}{\sqrt(SE(\hat{\beta}_{2010}[i]) + SE(\hat{\beta}_{2013}[i]))}.$$
-
-\subsection{Results}
-
-For the Chow statistic we find
-
-For the individual Z-tests we present the figure below in order of absolute value $Z_i$.
-
-Discussion
-
-\section{Investigating Question 5: Previous Earnings v Behavior}
+\section{Investigating Question 4: Previous Earnings v Behavior}
 \subsection{Experimental Setup}
@@
 Hence, we find the frequency of `above-above' vs `above-below' to be a good test for a behavioral factor.
-For each driver we output the tip statistic,
+For each driver we output the tip statistic
 
@@
 
+\section{Investigating Question 5: Average Tip Percentage vs Income}
+Next, we look at whether there is a relationship between passenger income and tip percentage. This involves several steps.
+\begin{itemize}
+\item We first create a grid of zip codes from querying latitude and longtitude for NY area based on Python's \texttt{geopy} package.
+\item Then we use the grid to match dropoff lat-long to zip code via MapReduce. Note that because the matrix is created already, zip code lookup on the grid take $O(1)$ time.
+\item We make the (crude) assumption that people who are getting dropped off are going home. We use the census data to match the zip code to median household income.
+\item We calculate the average tip percentage for that zip code to investigate the relationship between median household income for dropoff zip code and average tip percentage.
+\end{itemize}
+
+\includegraphics[scale=.4]{zip_income.png}
+As can be seen from the above chart, there seems to be a very small positive relationship between median household income and average tip percentage.
 %------------------------------------------------
@@
 
-\section*{Appendices}
+\section{Appendices}
 
@@
 \noindent https://s3.amazonaws.com/trip-fare-join/2010/part-0000X
-\\
 \noindent https://s3.amazonaws.com/trip-fare-join/2013/part-0000X

### 2015-05-11 Report/article_2.tex
Updated Report

@@
 We present our findings, issues encountered, and experimental setup.
-Of these findings, the most notable include: there are outlier drivers who are on average more productive per day, underperforming drivers generally do not improve (in relation to their peers), the sensitivity of fares to environment variables are seemingly invariant, drivers can expect better tips from cars with certain numbers of passengers, and previous tips affect driver behavior for the next immediate trip.
+Of these findings, the most notable include: there are outlier drivers who are on average more productive per day, the sensitivity of fares to environment variables are seemingly invariant, the distribution of actual to predicted fares implies a number of overcharges that should be investigated, drivers can expect better tips from passenger drop-off to certain areas, and previous tips affect driver behavior for the next trip.
 
@@
 
-All of the analyses above used the trips and fares data for NYC in 2010. Comparisons over time used the data for 2013 as well. This data is all publicly available. \cite{NYC_DATA}
+All of the analyses above used the trips and fares data for NYC in 2010. Comparisons over time used the data for 2013 as well. This data is all publicly available.
 
 As a starting point for our analyses we joined these two datasets.
-In order to perform this inner join, we used MrJob, a python tool for using Hadoop-Streaming with a programming model that more closely follows Java.\cite{MRJOB}  For instance, instead of writing a reducer that iterates through each line to collect values for a given key, MrJob provides the reducer with a key and list of values.
+In order to perform this inner join, we used MrJob, a python tool for using Hadoop-Streaming with a programming model that more closely follows Java.
+For instance, instead of writing a reducer that iterates through each line to collect values for a given key, MrJob provides the reducer with a key and list of values.
 
@@
 
-Additionally, we approximated the standard error of $\beta$.
+Additionally, we approximated the variance of $\hat{\beta}$.
 We found this to be an interesting problem as an exact result would require quadratic time, while an approximate one, like jackknifing would also be too computationally intensive (and of the same complexity though not assuming a functional form like homoskedasticity).
 We devised an approximation by randomly sampling 300,000 points (call this $\tilde{X}$) without replacement from our dataset.
-We then approximated $Var(\beta)$ by $\text{MSE}*(\frac{N}{300000}\tilde{X}^T\tilde{X})^{-1}$.
-We used Spark for this sampling procedure to produce our homeskedastic standard errors.
+We then approximated $Var(\hat{\beta})$ by $\text{MSE}*(\frac{N}{300000}\tilde{X}^T\tilde{X})^{-1}$.
+We used Spark for this sampling procedure to produce our homoskedastic standard errors.
+
+In order to test time invariance we fit regressions on both the 2010 and 2013 joined taxi data (separately and together).
+Using Spark, we jointly test whether the regression coefficients are the same as well as perform a test per coefficient for equality.
+
+The first analysis is conducted via a Chow Test.  The null hypothesis of the Chow Test is that $\beta_{2010} = \beta_{2013}$.  The Chow Statistic is formed by
+
+$$\frac{(SSR_{comb} - (SSR_{2010} + SSR_{2013}))/k}{(SSR_{2010} + SSR_{2013})/(N_{2010} + N_{2013}-2k)}$$
+
+\noindent which is distributed $F(k, N_{2010} + N_{2013}-2k)$. \cite{CHOW}
+
+We used Spark to calculate the Sum-of-Squared-Residual terms from the fitted models as well as the counts. This was included in the original regression code (so as to avoid loading the data twice).
+
+Secondly, we performed a Z-Test on each individual coefficient. Here
+
+$$ Z_i = \frac{\beta_{2010}[i] - \beta_{2013}[i]}{\sqrt(SE(\beta_{2010}[i]) + SE(\beta_{2013}[i]))}.$$
 \subsection{Results}
 
-The beta values are presented below in order of largest value to smallest value. A flag is included for significance.
+The top 10 most positive $\hat{\beta}$ values are presented in a table below.
 
-Below, we present them in order of most to least significant (in terms of the t-statistic). We include a flag for significance.
+\begin{tabular}{ |l|l| }
+  \hline
+  \multicolumn{2}{|c|}{10 Largest Positive} \\
+  \hline
+Distance &	5.50 \\
+ Month\_8  &	0.27\\
+ hour\_15 &	0.16\\
+ hour\_4 &	0.15\\
+ hour\_14 &	0.15\\
+ hour\_5 &	0.14\\
+ hour\_0 &	0.14\\
+ Thur &	0.11\\
+ Month\_9 &	0.11\\
+ Sat\_hour\_23 & 	0.09\\
+  \hline
+\end{tabular}
 
-Please see appendix for exact betas and standard errors.
+The top 10 most negative $\hat{\beta}$ values are presented in a table below. \\
 
-We see these as an interesting factor in driver behavior because...
+\begin{tabular}{ |l|l| }
+  \hline
+  \multicolumn{2}{|c|}{10 Smallest Negative} \\
+  \hline
+ Month\_0 &	-0.34\\
+ hour\_7 &	-0.31\\
+ Month\_1 &	-0.23\\
+ hour\_6 & 	-0.21\\
+ Month\_6 &	-0.17\\
+ hour\_20 &	-0.16\\
+ hour\_19 &	-0.16\\
+ Mon &	-0.14\\
+ hour\_10 &	-0.13\\
+ Month\_3 &	-0.12\\
+  \hline
+\end{tabular}\\
+
+The dummies all had very high standard errors while distance was highly significant.
+We suspect that the dummy standard errors would be lower without our approximation.
+Please see outputs for exact betas and standard errors.
+
+We find the above results interesting in terms of driver incentives.
+Those who work part-time should be partial to working time periods that align with high dummy values (these would lead to higher fares and tips for conceivably the same distances).
+They should avoid the low beta value periods.
+If the taxi company wants to smooth availability, assuming drivers already have some idea about these results, they should offer financial incentives by time period to even out the beta distribution.
+
+The Chow Test yielded that we cannot reject the null hypothesis that the sensitivities are jointly equal for the 2 different years.
+However, the distance beta Z-value would lead us to reject the hypothesis of equality for the sensitivity for the two different years.
+The other Z-values are not significant.
+Hence, the results of the Z-test and Chow test seem consistent.
 
@@
 
-%----------------------------------------------------------------------------------------
-%	REFERENCE LIST
-%----------------------------------------------------------------------------------------
-
-\begin{thebibliography}{99} % Bibliography - this is intentionally simple in this template
-
-\bibitem[Figueredo and Wolf, 2009]{Figueredo:2009dg}
-Figueredo, A.~J. and Wolf, P. S.~A. (2009).
-\newblock Assortative pairing and life history strategy - a cross-cultural
-  study.
-\newblock {\em Human Nature}, 20:317--330.
-
-\end{thebibliography}
+We have outlined a number of factors that affect driver behavior.
+Moreover, we provide a guide on how to use modern big data tools for this type of analysis.
+Each question could be greatly expanded into more parts.
+We offer this report as a basic framework to aid in such a further investigation.
 
@@
 
-Each component has a corresponding question folder with instructions as to how to run the code.
+\noindent Each component has a corresponding question folder with instructions as to how to run the code.
 We have excluded all of our AWS access information from the code and put in dummy values.
@@
 
-\noindent https://s3.amazonaws.com/trip-fare-join/2010/part-0000X
+\noindent https://s3.amazonaws.com/trip-fare-join/2010/part-0000X\\
 \noindent https://s3.amazonaws.com/trip-fare-join/2013/part-0000X
-\\
-
-\noindent where X ranges from 0 to 6.
+\\\\\
+where X ranges from 0 to 6.
 
 \subsection{Individual Contributions}
+\subsubsection*{Daniel}
+\begin{itemize}
+  \item Preprocessing Taxi Data Inputs
+  \item Question 1: Driver Productivity Differences
+  \item Question 2: Fare Sensitivities
+  \item Question 4: Previous Earnings v Behavior
+  \item Report - Writing
+  \item Code - Organization for Submission
+\end{itemize}
+
+\subsubsection*{Yoon}
+\begin{itemize}
+  \item Question 3: Driver Performance v Fare Sensitivities
+  \item Question 5: Average Tip Percentage vs Income
+  \item Presentation - Writing and Formatting
+  \item Report - Writing
+  \item Code - Organization for Submission
+\end{itemize}
 

### 2015-05-11 Q5 - Zip Tip/zip_tip.py
gave Zip Tip a Q number

@@ -1,0 +1,45 @@
+import sys
+import string
+import os
+import numpy as np
+import cPickle as pickle
+
+from mrjob.job import MRJob
+m = pickle.load(open("zip_grid.p","rb"))
+
+class ZipTip(MRJob):
+
+	def mapper(self, _, line):
+		def return_zip(lat, lon, nw_lat=40.917, nw_lon=-74.260, steps=0.002):
+		    try:
+		    	lat, lon = float(lat), float(lon)     
+		    	i, j = int((nw_lat - lat)/steps), int((lon - nw_lon)/steps)
+		        return m[i][j]
+		    except:
+		        return -1
+		#take all csv objects
+		items = line.strip('"').split('"\t"')
+		items = items[0].split(',') + items[1].split(',')
+
+		lon,lat = tuple([x for x in items[12:14]])
+		#some seem to be transposed
+		zip_c = max(return_zip(lat,lon), return_zip(lon,lat))
+		tips = float(items[-3])
+		#all - tip
+		rev = float(items[-1]) - tips
+		try:
+			tip_per = float(tips)/rev 
+		except:
+			tip_per = 0
+		if tip_per > 0:
+			yield (str(zip_c), float(tip_per))
+
+
+	def reducer(self, key, values):
+		tips = [v for v in values]
+		count = len(tips)
+		avg_tip = sum(tips)/float(len(tips))
+		yield (key, [avg_tip, count])
+
+if __name__ == '__main__':
+    ZipTip.run()

### 2015-05-11 QX - Reg Errors/histos.py
done?

@@ -1,50 +1,0 @@
-import glob
-import numpy as np
-import matplotlib.pyplot as plt
-import matplotlib.mlab as mlab
-import json
-
-f_2010 = glob.glob('outputs/2010/p*')
-
-f_2013 = glob.glob('outputs/2013/p*')
-
-def file_loop(f):
-    lines = [line.strip().split('\t') for line in open(f)]
-    vals = [(l[0],json.loads(l[1])[0],json.loads(l[1])[1] )for l in lines]
-    return vals
-
-v_2010 = []
-for f in f_2010:
-    v_2010.extend(file_loop(f))
-
-
-v_2013 = []
-for f in f_2013:
-    v_2013.extend(file_loop(f))
-
-def make_graph(values, label):
-	n, bins, patches = plt.hist(values, 35, normed=1, stacked=1,facecolor='green', alpha=0.75)
-	# add a 'best fit' line (assuming normality)
-	y = mlab.normpdf( bins, np.mean(values), np.std(values))
-	l = plt.plot(bins, y, 'r--', linewidth=1)
-
-	plt.xlabel(label)
-	plt.ylabel('Scaled Frequency')
-	plt.title(r'Distribition of ?' + label)
-	#plt.axis([.1, .9,0,10])
-	plt.grid(True)
-
-	plt.savefig(label + '.png', dpi=400)
-	plt.close()
-
-#strip teh very strange outliers
-avg_err_10 = [v[1] for v in v_2010 if (x<1 and x>-1)]
-abs_err_10 = [v[2] for v in v_2010 if (x<50 and x>50)]
-
-avg_err_13 = [v[1] for v in v_2013 if (x<1 and x>-1)]
-abs_err_13 = [v[2] for v in v_2010 if (x<50 and x>50)]
-
-make_graph(avg_err_10, 'Average Errors 2010')
-make_graph(abs_err_10, 'Absolute Errors 2010')
-make_graph(avg_err_13, 'Average Errors 2013')
-make_graph(avg_err_13, 'Absolute Errors 2013')

### 2015-05-11 QX - Reg Errors/reg_errors.py

@@ -1,36 +1,0 @@
-import sys
-import string
-import os
-
-
-from mrjob.job import MRJob
-
-
-class RegErrors(MRJob):
-
-	def mapper(self, _, line):
-		x = line.strip('"(').split(",")[:4]
-		actual,pred = [float(a) for a in x[:2]]
-		m1 = x[2].strip('"[').strip('" [').strip('\'')
-		m2 = x[3].strip('"[').strip('" [').strip('\'')
-		driver = m1 + "_" + m2
-		yield (driver, [actual,pred])
-
-
-	def reducer(self, key, values):
-		vals = [v for v in values]
-		percs = []
-		absos = []
-		for v in vals:
-			try:
-				percs.append(float(v[0] - v[1]) / v[1])
-			except:
-				pass
-			try:
-				absos.append(float(v[0] - v[1]))
-			except:
-				pass
-		yield (key, [float(sum(percs))/len(percs), float(sum(absos))/len(absos)])
-
-if __name__ == '__main__':
-    RegErrors.run()

### 2015-05-11 QX - Zip Tip/zip_tip.py

@@ -1,45 +1,0 @@
-import sys
-import string
-import os
-import numpy as np
-import cPickle as pickle
-
-from mrjob.job import MRJob
-m = pickle.load(open("zip_grid.p","rb"))
-
-class ZipTip(MRJob):
-
-	def mapper(self, _, line):
-		def return_zip(lat, lon, nw_lat=40.917, nw_lon=-74.260, steps=0.002):
-		    try:
-		    	lat, lon = float(lat), float(lon)     
-		    	i, j = int((nw_lat - lat)/steps), int((lon - nw_lon)/steps)
-		        return m[i][j]
-		    except:
-		        return -1
-		#take all csv objects
-		items = line.strip('"').split('"\t"')
-		items = items[0].split(',') + items[1].split(',')
-
-		lon,lat = tuple([x for x in items[12:14]])
-		#some seem to be transposed
-		zip_c = max(return_zip(lat,lon), return_zip(lon,lat))
-		tips = float(items[-3])
-		#all - tip
-		rev = float(items[-1]) - tips
-		try:
-			tip_per = float(tips)/rev 
-		except:
-			tip_per = 0
-		if tip_per > 0:
-			yield (str(zip_c), float(tip_per))
-
-
-	def reducer(self, key, values):
-		tips = [v for v in values]
-		count = len(tips)
-		avg_tip = sum(tips)/float(len(tips))
-		yield (key, [avg_tip, count])
-
-if __name__ == '__main__':
-    ZipTip.run()

### 2015-05-11 Report/article_2.tex
straggling citation
final

@@
 \noindent Using all data on NYC taxi trips and fares from 2010 and 2013 we perform a series of studies to understand motivating factors behind driver behavior as well as outcomes from those incentives.
-In order to analyze this large dataset ($\approx$14GB), we employ a number of big data processing technologies that work on the Hadoop framework: Hive, Spark, and MapReduce programs.
+In order to analyze this large dataset ($\approx$14GB), we employ a number of big data processing technologies that work on the Hadoop framework: Spark and MapReduce programs.
 We present our findings, issues encountered, and experimental setup.
@@
 
-\noindent which is distributed $F(k, N_{2010} + N_{2013}-2k)$. \cite{CHOW}
+\noindent which is distributed $F(k, N_{2010} + N_{2013}-2k)$.
 

### 2015-05-11 Q7-tip/m.py
q7-tip

@@ -1,0 +1,17 @@
+#!/usr/bin/python
+
+import sys
+import string
+
+#map passenger_count and credit card payment_amount and tip_amount
+for line in sys.stdin:
+	try:
+		items = line.strip('"').split('"\t"')
+		item = items[0].split(',') + items[1].split(',')
+		if len(item) == 21:
+			key = item[7]
+			if item[14].upper() == 'CRE':
+				values = item[15]+ ',' + item[18]
+				print '%s,%s' %(key, values)
+	except:
+		continue

### 2015-05-11 Q7-tip/mct.py

@@ -1,0 +1,19 @@
+#!/usr/bin/python
+
+import sys
+import string
+
+#map payment_type, count(1) for each positive tip
+for line in sys.stdin:
+	try:
+		items = line.strip().split(",")
+		if items[0] == 'medallion':
+			continue
+		key = items[4].upper()
+		if float(items[8]) > 0:
+			values = 1
+		else:
+			values = 0
+		print '%s,%s' %(key, values)
+	except:
+		pass

### 2015-05-11 Q7-tip/mmed.py

@@ -1,0 +1,16 @@
+#!/usr/bin/python
+
+import sys
+import string
+from datetime import datetime
+
+fmt = '%Y-%m-%d %H:%M:%S'
+for line in sys.stdin:
+	try:
+		items = line.split(',')
+		weekday = datetime.strptime((items[3]), fmt).strftime("%A")
+		key = weekday
+		values = items[0]
+		print '%s,%s' %(key, values)
+	except:
+		continue

### 2015-05-11 Q7-tip/mn0.py

@@ -1,0 +1,16 @@
+#!/usr/bin/python
+
+import sys
+import string
+
+for line in sys.stdin:
+	try:
+		items = line.strip().split(",")
+		if items[0] == 'medallion':
+			continue
+		key = items[4].upper()
+		if float(items[8]) > 0:
+			values = items[5] + "," + items[8]
+			print '%s,%s' %(key, values)
+	except:
+		pass

### 2015-05-11 Q7-tip/pack/plot.py

@@ -1,0 +1,81 @@
+import matplotlib.pyplot as plt
+import pandas as pd
+import numpy as np
+
+#passenger plot
+passenger = pd.DataFrame(pd.read_csv('pass.csv'))
+plot = passenger.plot(kind='bar', x='pass',  y='pctip')
+plt.title('Credit Card Tip Percentage vs Passenger Count')
+plt.xlabel('Number of Passengers')
+plt.ylabel('Percentage')
+plt.savefig('tip_passenger.pdf')
+plt.show()
+
+a = pd.read_csv("pc0.csv")
+b = pd.read_csv("pyno0.csv")
+ct = pd.read_csv("pyct.csv")
+
+#ct
+ct.plot(kind='line', x='type', y='pcct', label = 'Tipped Rides/Total Rides', color='black')
+plt.xlabel('Payment Type')
+plt.ylabel('Percentage')
+plt.legend(loc = 'upper right')
+plt.title('Ratio of Tipped Rides vs Total Rides by Payment Type')
+plt.savefig('pct.pdf')
+plt.show()
+
+#0, no0
+merged = a.merge(ct, on='type')
+merged.to_csv("pcmerge.csv", index=False)
+pcm = pd.read_csv('pcmerge.csv')
+pcm.plot(kind='line', x = 'type', y='pcct', label='Tipped Rides/Total Rides', color = 'black')
+pcm.plot(kind='line', x='type', y='pc', label='Tip By Type', color ='b')
+plt.xlabel('Payment Type')
+plt.ylabel('Percentage')
+plt.legend(loc = 'upper right')
+plt.savefig('pct_pc0.pdf')
+plt.show()
+
+#ct, 0, no0
+mct = pcm.merge(b, on='type')
+mct.to_csv("pctmerge.csv", index=False)
+mctp = pd.read_csv("pctmerge.csv")
+mctp.plot(kind='line', x='type', y='pcct', label = 'Tipped Rides/Total Rides', color = 'black')
+mctp.plot(kind='line', x = 'type', y='pc', label = 'Tip By Type', color = 'b')
+mctp.plot(kind='line', x='type', y='pcno0', label = 'Tip By Type (Removed 0s)', color = 'g')
+plt.xlabel('Payment Type')
+plt.ylabel('Percentage')
+plt.legend(loc = 'upper right')
[…]

### 2015-05-11 Q7-tip/r0.py

@@ -1,0 +1,32 @@
+#!/usr/bin/python
+
+import sys
+
+current_key = None
+current_fare = None
+current_tip = None
+
+for line in sys.stdin:
+    
+	key, fare, tip = line.split(",")
+	try:
+		fare = float(fare)
+		tip = float(tip)
+	except:
+		continue
+	#if same key, add fare, add tip
+	if key == current_key:
+		current_fare += fare
+		current_tip += tip
+	else:
+	#not 1st line && diff key -> print current entry, then set current to new current
+		if current_key:
+			pctip = 100*(current_tip / current_fare)
+			print "%s,%s" %( current_key, pctip)
+		current_key = key
+		current_fare = fare
+		current_tip = tip
+        
+#print final line
+pctip = 100*(current_tip / current_fare)
+print "%s,%s" %( current_key, pctip )

### 2015-05-11 Q7-tip/rct.py

@@ -1,0 +1,31 @@
+#!/usr/bin/python
+
+import sys
+
+current_key = None
+current_value = 1
+total_value = 1
+
+for line in sys.stdin:
+    
+	key, value = line.split(",")
+	try:
+		value = float(value)
+	except:
+		continue
+	#if same key, add fare, add tip
+	if key == current_key:
+		current_value = current_value + value
+		total_value = total_value + 1
+	else:
+	#not 1st line && diff key -> print current entry, then set current to new current
+		if current_key:
+			pctip = 100*(current_value / total_value)
+			print "%s,%s" %( current_key, pctip)
+		current_key = key
+		current_value = value
+		total_value = 1
+        
+#print final line
+pctip = 100*(current_value / total_value)
+print "%s,%s" %( current_key, pctip )

### 2015-05-11 Q6-tip/m0.py
renumber to Q6

@@ -1,0 +1,16 @@
+#!/usr/bin/python
+
+import sys
+import string
+
+#map payment_type, fare_amount, tip_amount
+for line in sys.stdin:
+	try:
+		items = line.strip().split(",")
+		if items[0] == 'medallion':
+			continue
+		key = items[4].upper()
+		values = items[5] + "," + items[8]
+		print '%s,%s' %(key, values)
+	except:
+		continue

### 2015-05-11 Q6-tip/mdate.py

@@ -1,0 +1,19 @@
+#!/usr/bin/python
+
+import sys
+import string
+from datetime import datetime
+
+#map day, tip_amount
+fmt = '%Y-%m-%d %H:%M:%S'
+for line in sys.stdin:
+	try:
+		items = line.strip('"').split('"\t"')
+		items = items[0].split(',') + items[1].split(',')
+		weekday = datetime.strptime((items[3]), fmt).strftime("%A")
+		if len(items) == 21:
+			key = weekday
+			values = items[18]
+			print '%s,%s' %(key, values)
+	except:
+		continue

### 2015-05-11 Q6-tip/pack/test.py

@@ -1,0 +1,37 @@
+import matplotlib.pyplot as plt
+import pandas as pd
+import numpy as np
+
+taxi = pd.read_csv('pyutaxi.csv')
+tip = pd.read_csv('pytipday.csv')
+tiptaxi = tip.merge(taxi, on='day')
+tiptaxi.to_csv('tiptaxi.csv', index=False)
+tiptaxi = pd.DataFrame(pd.read_csv('tiptaxi.csv'))
+tiptaxi['ratio'] = tiptaxi['tip']/tiptaxi['taxi']
+print tiptaxi
+tiptaxi.to_csv('tiptaxi.csv', index=False)
+
+tiptaxi.plot(kind='line', x='day', y='ratio')
+plt.title('Tips Per Taxi By Day')
+plt.xlabel('Day')
+plt.ylabel('Tips Earned ($)')
+plt.savefig('tips_per_taxi.pdf')
+plt.show()
+# f=open('tiptaxi.csv','r')
+# line = f.readlines
+# for l in line:
+#     day, ntip, ntaxi = line.strip().split(",")
+#     ntip = float(ntip)
+#     ntaxi = float(ntaxi)
+#     ratio = ntip/ntaxi
+#     print "%s,%d" %(day, ratio)
+    
+# tiptaxi = pd.read_csv('tiptaxi.csv')
+# ratio = tiptaxi[1]/tiptaxi[3]
+# print ratio
+# for line in tiptaxi:
+#     day, ntip, index, ntaxi = line.strip().split(",")
+#     ntip = float(ntip)
+#     ntaxi = float(ntaxi)
+#     ratio = ntip/ntaxi
+#     print "%s,%d" %(day, ratio)

### 2015-05-11 Q6-tip/r.py

@@ -1,0 +1,29 @@
+#!/usr/bin/python
+
+import sys
+
+current_key = None
+current_fare = None
+current_tip = None
+
+for line in sys.stdin:
+    
+	key, fare, tip = line.strip().split(",")
+	fare = float(fare)
+	tip = float(tip)
+	#if same key, add fare, add tip
+	if key == current_key:
+		current_fare += fare
+		current_tip += tip
+	else:
+	#not 1st line && diff key -> print current entry, then set current to new current
+		if current_key:
+			pctip = 100*(current_tip / current_fare)
+			print "%s,%s" %( current_key, pctip)
+		current_key = key
+		current_fare = fare
+		current_tip = tip
+        
+#print final line
+pctip = 100*(current_tip / current_fare)
+print "%s,%s" %( current_key, pctip )

### 2015-05-11 Q6-tip/rdate.py

@@ -1,0 +1,26 @@
+#!/usr/bin/python
+
+import sys
+
+current_key = None
+current_tip = None
+
+for line in sys.stdin:
+    
+	key, tip = line.strip().split(",")
+	try:
+		tip = float(tip)
+	except ValueError:
+		continue
+	#if same key, add fare, add tip
+	if key == current_key:
+		current_tip += tip
+	else:
+	#not 1st line && diff key -> print current entry, then set current to new current
+		if current_key:
+			print "%s,%s" %( current_key, current_tip)
+		current_key = key
+		current_tip = tip
+        
+#print final line
+print "%s,%s" %( current_key, current_tip )

### 2015-05-11 Q6-tip/rmed.py

@@ -1,0 +1,27 @@
+#!/usr/bin/python
+
+import sys
+
+current_key = None
+current_med = None
+
+for line in sys.stdin:
+    
+	key, med = line.strip().split(",")
+	#if same key, add fare, add tip
+	if key == current_key:
+		if med != current_med:
+			sum = sum + 1
+			current_med = med
+		else:
+			continue
+	else:
+	#not 1st line && diff key -> print current entry, then set current to new current
+		if current_key:
+			print "%s,%s" %( current_key, sum)
+		current_key = key
+		current_med = med
+		sum = 1
+        
+#print final line
+print "%s,%s" %( current_key, sum )

### 2015-05-11 Q7-tip/m.py
fixed files/names

@@ -1,17 +1,0 @@
-#!/usr/bin/python
-
-import sys
-import string
-
-#map passenger_count and credit card payment_amount and tip_amount
-for line in sys.stdin:
-	try:
-		items = line.strip('"').split('"\t"')
-		item = items[0].split(',') + items[1].split(',')
-		if len(item) == 21:
-			key = item[7]
-			if item[14].upper() == 'CRE':
-				values = item[15]+ ',' + item[18]
-				print '%s,%s' %(key, values)
-	except:
-		continue

### 2015-05-11 Q7-tip/mdate.py

@@ -1,19 +1,0 @@
-#!/usr/bin/python
-
-import sys
-import string
-from datetime import datetime
-
-#map day, tip_amount
-fmt = '%Y-%m-%d %H:%M:%S'
-for line in sys.stdin:
-	try:
-		items = line.strip('"').split('"\t"')
-		items = items[0].split(',') + items[1].split(',')
-		weekday = datetime.strptime((items[3]), fmt).strftime("%A")
-		if len(items) == 21:
-			key = weekday
-			values = items[18]
-			print '%s,%s' %(key, values)
-	except:
-		continue

### 2015-05-11 Q7-tip/pack/plot.py

@@ -1,81 +1,0 @@
-import matplotlib.pyplot as plt
-import pandas as pd
-import numpy as np
-
-#passenger plot
-passenger = pd.DataFrame(pd.read_csv('pass.csv'))
-plot = passenger.plot(kind='bar', x='pass',  y='pctip')
-plt.title('Credit Card Tip Percentage vs Passenger Count')
-plt.xlabel('Number of Passengers')
-plt.ylabel('Percentage')
-plt.savefig('tip_passenger.pdf')
-plt.show()
-
-a = pd.read_csv("pc0.csv")
-b = pd.read_csv("pyno0.csv")
-ct = pd.read_csv("pyct.csv")
-
-#ct
-ct.plot(kind='line', x='type', y='pcct', label = 'Tipped Rides/Total Rides', color='black')
-plt.xlabel('Payment Type')
-plt.ylabel('Percentage')
-plt.legend(loc = 'upper right')
-plt.title('Ratio of Tipped Rides vs Total Rides by Payment Type')
-plt.savefig('pct.pdf')
-plt.show()
-
-#0, no0
-merged = a.merge(ct, on='type')
-merged.to_csv("pcmerge.csv", index=False)
-pcm = pd.read_csv('pcmerge.csv')
-pcm.plot(kind='line', x = 'type', y='pcct', label='Tipped Rides/Total Rides', color = 'black')
-pcm.plot(kind='line', x='type', y='pc', label='Tip By Type', color ='b')
-plt.xlabel('Payment Type')
-plt.ylabel('Percentage')
-plt.legend(loc = 'upper right')
-plt.savefig('pct_pc0.pdf')
-plt.show()
-
-#ct, 0, no0
-mct = pcm.merge(b, on='type')
-mct.to_csv("pctmerge.csv", index=False)
-mctp = pd.read_csv("pctmerge.csv")
-mctp.plot(kind='line', x='type', y='pcct', label = 'Tipped Rides/Total Rides', color = 'black')
-mctp.plot(kind='line', x = 'type', y='pc', label = 'Tip By Type', color = 'b')
-mctp.plot(kind='line', x='type', y='pcno0', label = 'Tip By Type (Removed 0s)', color = 'g')
-plt.xlabel('Payment Type')
-plt.ylabel('Percentage')
-plt.legend(loc = 'upper right')
[…]

### 2015-05-11 Q7-tip/pack/test.py

@@ -1,37 +1,0 @@
-import matplotlib.pyplot as plt
-import pandas as pd
-import numpy as np
-
-taxi = pd.read_csv('pyutaxi.csv')
-tip = pd.read_csv('pytipday.csv')
-tiptaxi = tip.merge(taxi, on='day')
-tiptaxi.to_csv('tiptaxi.csv', index=False)
-tiptaxi = pd.DataFrame(pd.read_csv('tiptaxi.csv'))
-tiptaxi['ratio'] = tiptaxi['tip']/tiptaxi['taxi']
-print tiptaxi
-tiptaxi.to_csv('tiptaxi.csv', index=False)
-
-tiptaxi.plot(kind='line', x='day', y='ratio')
-plt.title('Tips Per Taxi By Day')
-plt.xlabel('Day')
-plt.ylabel('Tips Earned ($)')
-plt.savefig('tips_per_taxi.pdf')
-plt.show()
-# f=open('tiptaxi.csv','r')
-# line = f.readlines
-# for l in line:
-#     day, ntip, ntaxi = line.strip().split(",")
-#     ntip = float(ntip)
-#     ntaxi = float(ntaxi)
-#     ratio = ntip/ntaxi
-#     print "%s,%d" %(day, ratio)
-    
-# tiptaxi = pd.read_csv('tiptaxi.csv')
-# ratio = tiptaxi[1]/tiptaxi[3]
-# print ratio
-# for line in tiptaxi:
-#     day, ntip, index, ntaxi = line.strip().split(",")
-#     ntip = float(ntip)
-#     ntaxi = float(ntaxi)
-#     ratio = ntip/ntaxi
-#     print "%s,%d" %(day, ratio)

### 2015-05-11 Q7-tip/r.py

@@ -1,29 +1,0 @@
-#!/usr/bin/python
-
-import sys
-
-current_key = None
-current_fare = None
-current_tip = None
-
-for line in sys.stdin:
-    
-	key, fare, tip = line.strip().split(",")
-	fare = float(fare)
-	tip = float(tip)
-	#if same key, add fare, add tip
-	if key == current_key:
-		current_fare += fare
-		current_tip += tip
-	else:
-	#not 1st line && diff key -> print current entry, then set current to new current
-		if current_key:
-			pctip = 100*(current_tip / current_fare)
-			print "%s,%s" %( current_key, pctip)
-		current_key = key
-		current_fare = fare
-		current_tip = tip
-        
-#print final line
-pctip = 100*(current_tip / current_fare)
-print "%s,%s" %( current_key, pctip )

### 2015-05-11 Q7-tip/rmed.py

@@ -1,27 +1,0 @@
-#!/usr/bin/python
-
-import sys
-
-current_key = None
-current_med = None
-
-for line in sys.stdin:
-    
-	key, med = line.strip().split(",")
-	#if same key, add fare, add tip
-	if key == current_key:
-		if med != current_med:
-			sum = sum + 1
-			current_med = med
-		else:
-			continue
-	else:
-	#not 1st line && diff key -> print current entry, then set current to new current
-		if current_key:
-			print "%s,%s" %( current_key, sum)
-		current_key = key
-		current_med = med
-		sum = 1
-        
-#print final line
-print "%s,%s" %( current_key, sum )

### 2015-05-11 Report/article_2.tex
edited report

@@
 \normalsize New York University \\ % Your institution
-\normalsize \href{mailto:ssz8@cornell.edu}{ssz8@cornell.edu} % Your email address
+\normalsize \href{mailto:ssz225@nyu.edu}{ssz225@nyu.edu} % Your email address
 \vspace{-5mm}
@@
 %----------------------------------------------------------------------------------------
-
 \begin{abstract}
-
 \noindent Using all data on NYC taxi trips and fares from 2010 and 2013 we perform a series of studies to understand motivating factors behind driver behavior as well as outcomes from those incentives.
@@
 We present our findings, issues encountered, and experimental setup.
-Of these findings, the most notable include: there are outlier drivers who are on average more productive per day, the sensitivity of fares to environment variables are seemingly invariant, the distribution of actual to predicted fares implies a number of overcharges that should be investigated, drivers can expect better tips from passenger drop-off to certain areas, and previous tips affect driver behavior for the next trip.
+Of these findings, the most notable include: tip aggregates and operating taxis are sensitive to day of the week, there are outlier drivers who are on average more productive per day, the sensitivity of fares to environment variables are seemingly invariant, the distribution of actual to predicted fares implies a number of overcharges that should be investigated, drivers can expect better tips from passenger drop-off to certain areas, and previous tips affect driver behavior for the next trip.
 
@@
 \item How is tip affected by the income of the passengers, as measured by the median household income of the zip code in which the passenger is dropped off?
-
+\item Do passengers tip more with respect to methods of payment and number of riders? Is driver behavior affected by tipping trends?
 \end{enumerate}
@@
 Mapper numbers were set to defaults.
-
 %------------------------------------------------
-
 \section{Investigating Question 1: Driver Productivity Differences}
@@
 \subsection{Results}
-\includegraphics[scale=.35]{Minutes_v_Rev.png}
 
@@
 It would appear they have much downtime during the day.
+\includegraphics[scale=.35]{Minutes_v_Rev.png}
 
@@
 Based on the regression output from the previous section, we plot the residual (i.e. actual fare minus predicted fare) for each trip and bin them into histograms.  
-\includegraphics[scale=.35]{figure_1.png}
+\includegraphics[scale=.34]{figure_1.png}
 While much the error may be coming from the fact that our regression model is very simple (for example, it doesn't take into account the fact that a distance of 1 mile means very different things in midtown versus in suburbs of Queens), there does seem to be a cohort of trips in which the customer is overcharged.
@@
 While the relationship is far from clear, there does seem to be a slight negative relationship between tip percentage and residual.
-\includegraphics[scale=.35]{figure_2.png}
+\includegraphics[scale=.345]{figure_2.png}
 
@@
 
-\includegraphics[scale=.4]{zip_income.png}
+\includegraphics[scale=.35]{zip_income.png}
 As can be seen from the above chart, there seems to be a very small positive relationship between median household income and average tip percentage.
@@
 
+\section{Investigating Question 6: Aggregate Trends in Tipping and Driver Behavior}
+\subsection{Experimental Setup}
+A series of MapReduce functions were utilized to generate aggregate trend visualizations of fares and the combined trips and fares data from 2010.
+Tipping and driver behaviors were examined using the following variables: number of passengers per trip, day of the week, and method of fare payment.
+
+\subsection{Results}
+To obtain an overview of tipping trends, the ratio of the count of rides with positive tips over all rides was plotted against the methods of payment.
+Percentage tip was calculated over all data as well as over data with tip equalling zero dropped.
+\includegraphics[scale=.35]{pct_pc0_pcno0.pdf}
+
+The graph points to a nearly absolute drop in the frequency of tipped trips when the payment method is not under automated reporting systems (i.e.
+credit cards).
+Filtering out zeros in tip records, the plot shows a fairly consistent rate of 20 percent in tips which agrees with the more reliably reported complete credit cards tip data.
+Therefore, it is unlikely that the drastic drop in frequency of tipped trips is due to passenger behavior changed based on payment type.
+This highlights the problem of drivers "cheating" on reporting approximately 96 percent of instances of non-credit card tips, making non-credit card tip data unreliable in raw form.
+This problem may be an issue for taxi companies when calculating compensation packages for contracted drivers.
+
+Since non-credit card tip data was determined to be unreliable, the plot of number of passengers against tip percentage was generated over only trips paid with credit cards.
+\includegraphics[scale=.35]{tip_passenger.pdf}
+Other than outliers of 0, 7, and 9 passengers, the tip percentage is fairly consistent at just slightly below 20 percent, which is in agreement with the aggregate statistic in the previous graph.
+Therefore, number of passengers is not supported as a source of deviation from normal tipping behavior.
+Overall, passenger tipping behavior is determined to be inelastic to payment type and number of taxi riders.
+
+To examine the effects of day of the week on tip and driver behavior, the total sum of tips and unique taxis were plotted against days.
+\includegraphics[scale=.32]{total_tip.pdf}
+\includegraphics[scale=.32]{total_taxis.pdf}
+Both graphs peak on Thursday and drop sharply on the weekends.
+With the exception of Monday and Tuesday, the graphs demonstrate similar trend shape.
+While other variables may be in play at the micro scale, the aggregate view supports a positive correlation between higher tips and willingness of drivers to participate in work.
+
+%------------------------------------------------
 \section{Concluding Remarks}
@@
 
+\subsubsection*{Sylvia}
+\begin{itemize}
+  \item Question 6: Aggregate Trends in Tipping and Driver Behavior
+  \item Presentation - Writing
+  \item Report - Writing
+  \item Code - Organization for Submission
+\end{itemize}
+
 \end{multicols}
